Supplementary Material for ``Evaluating infectious disease forecasts with allocation scoring rules''

### 2023-12-12 alloscore_manuscript/supplement.rnw
add quantile fig, change error measure in sec:numeric

For future reference, we record several key properties of quantile functions.

The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ form a discrete subset of $(0,1)$ and correspond to
the non-zero width intervals $[q_{\alpha_i,F}^{-},q_{\alpha_i,F}^{+}]$ where $F$ is constant with values 
$\{\alpha_i\}$.  Conversely, if $F$ is strictly increasing on a (Borel) set $A\subset \mathbb{R}$, then the restriction
$Q_{F}\big|_{F(A)}=F^{-1}\big|_{F(A)}$ of $Q_F$ to $F(A)$ is in fact a real-valued function which is left-continuous
and provides an inverse to $F$ on $F(A)$. If the the support $\mathrm{supp}(F) = \{x \mid 0<F(x)<1\}$ of $F$ is such an
$A$, then extending $F^{-1}$ by $F^{-1}(0)=\inf(\mathrm{supp}(F))$ and $F^{-1}(1)=\sup(\mathrm{supp}(F))$ provides an
inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$, so that
$\{\alpha \mid F(x-)< \alpha < F(x)\}$ is a non-empty set disjoint from $F(A)$, then $Q_F$ takes the constant value $x$ 
on the closure $\{\alpha \mid F(x-)\leq \alpha \leq F(x)\}$. Conversely, if $F$ has no discrete component on $A$, 
then $Q_{F}\big|_{F(A)}$ is strictly increasing.

Moreover, it can be said that $Q_F$ is increasing as a set-valued function on $\mathbb{R}$ in the generalized sense that 
$(q_{\alpha, F} - q_{\beta, F})(\alpha-\beta) \geq 0$ whenever
$q_{\alpha, F} \in Q_F(\alpha)$ and $q_{\beta, F} \in Q_F(\beta)$, that is, the set 
$\mathrm{graph}(Q_F) = \{(\alpha, q) \mid q \in Q_F(\alpha)\}$ has no downward sloping secants. Conversely, given such an 
increasing set-valued function $Q$ on $(0,1)$, we can construct a right-continuous function $F_Q$ from $\mathbb{R}$ to $[0,1]$ 
which will be the cdf of the random variable $Y_Q := \min(Q(U))$ where $U \sim \mathrm{Unif}[0,1]$.

@@

Exact computation of the allocation score $S_A(F, y; K)$ of a forecast $F$ requires an explicit choice of and expression for a 
solution $x^{F,K} = \bm{q}_{\tau^{\star},F} \in \undertilde{\bm{F}}^{-1}(\tau^{\star})$ to \eqref{eqn:quantiles-sum-to-K} where
$\tau^{\star}$ solves \eqref{eqn:tau-inclusion}. For practical scenarios in which the forecast $F$ has a form preventing 
analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we seek a method of generating an approximation $\tilde{\tau}$ to 
a solution $\tau^{\star}$ which offers control (at least in principle) of the Bayes risk error or ``optimality gap''
\begin{align}
\left|\mathbb{E}_F\left[s_A(\bm{q}_{\tilde{\tau},F}, Y)\right] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right| = 
L\left| \sum_{i=1}^{N}\mathbb{E}_{F_i}\left[(Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+\right]\right|
\end{align}
where $\bm{q}_{\tilde{\tau},F}$ solves the linear program
$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tilde{\tau})}x_i = K$.

>>>>

For future reference, we record several key properties of quantile functions. An illuminating source for these
facts is \cite{rockafellar2014random}, from which figure \ref{fig:FQ} is adapted.

The probability levels $\{\alpha_i\}$ for which $\#Q_F(\alpha_i)>1$ ($\{\alpha_3\}$ in the example of figure \ref{fig:FQ}) form a 
discrete subset of $(0,1)$ and correspond to
the non-zero width intervals $[q_{\alpha_i,F}^{-},q_{\alpha_i,F}^{+}]$ where $F$ is constant with values 
$\{\alpha_i\}$.  Conversely, if $F$ is strictly increasing on a (Borel) set $A\subset \mathbb{R}$, then the restriction
$Q_{F}\big|_{F(A)}=F^{-1}\big|_{F(A)}$ of $Q_F$ to $F(A)$ is in fact a real-valued function which is left-continuous
and provides an inverse to $F$ on $F(A)$. If the the support $\mathrm{supp}(F) = \{x \mid 0<F(x)<1\}$ of $F$ is such an
$A$, then extending $F^{-1}$ by $F^{-1}(0)=\inf(\mathrm{supp}(F))$ and $F^{-1}(1)=\sup(\mathrm{supp}(F))$ provides an
inverse to $F$ on $F(\mathbb{R})\subset [0,1]$. If $F$ has a point mass at $x\in A$ (e.g., $x_1 \in [0,\infty)$ in figure \ref{fig:FQ}),
so that $\{\alpha \mid F(x-)< \alpha < F(x)\}$ (e.g., $(\alpha_1, \alpha_2)$ in figure \ref{fig:FQ}) is a non-empty set disjoint from $F(A)$,
then $Q_F$ takes the constant value $x$ 
on the closure $\{\alpha \mid F(x-)\leq \alpha \leq F(x)\}$. Conversely, if $F$ has no discrete component on $A$, 
then $Q_{F}\big|_{F(A)}$ is strictly increasing (e.g., on $[\alpha_2, \infty)$ in figure \ref{fig:FQ}).

Moreover, it can be said that $Q_F$ is increasing as a set-valued function on $\mathbb{R}$ in the generalized sense that 
$(q_{\alpha, F} - q_{\beta, F})(\alpha-\beta) \geq 0$ whenever
$q_{\alpha, F} \in Q_F(\alpha)$ and $q_{\beta, F} \in Q_F(\beta)$, that is, the set 
$\mathrm{graph}(Q_F) = \{(\alpha, q) \mid q \in Q_F(\alpha)\}$ has no downward sloping secants. Conversely, given such an 
increasing set-valued function $Q$ on $(0,1)$, we can construct a right-continuous function $F_Q$ from $\mathbb{R}$ to $[0,1]$ 
which will be the cdf of the random variable $Y_Q := \min(Q(U))$ where $U \sim \mathrm{Unif}[0,1]$.
\begin{figure}
\vspace{-4cm}
\includegraphics[width=.9\textwidth]{../figures/FQ.pdf}
\vspace{-15mm}
\caption{Example of a cdf $F$ (with graph including solid line and filled points) superimposed on its 
quantile function $Q_F$ (with graph including dashed and solid lines as well as both filled and empty points).}
\label{fig:FQ}
\end{figure}

@@

Exact computation of the allocation score $S_A(F, y; K)$ of a forecast $F$ requires an explicit choice of and expression for a 
solution $x^{F,K} = \bm{q}_{\tau^{\star},F} \in \undertilde{\bm{F}}^{-1}(\tau^{\star})$ to \eqref{eqn:quantiles-sum-to-K} where
$\tau^{\star}$ solves \eqref{eqn:tau-inclusion}. For practical scenarios in which the forecast $F$ has a form preventing 
analytical solutions to \eqref{eqn:quantiles-sum-to-K}, we seek a method of generating an approximation $\tilde{\tau}$ to 
a solution $\tau^{\star}$ which offers control (at least in principle) over a robust measure of scoring rule error such as
the sum of absolute coordinate errors
\begin{align}
\mathrm{SAE}(Y,\tilde{\tau}) = \sum_{i=1}^{N} \left| (Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+ \right|
\end{align}
% of the Bayes risk error or ``optimality gap''
% \begin{align}
% \left|\mathbb{E}_F\left[s_A(\bm{q}_{\tilde{\tau},F}, Y)\right] - \mathbb{E}_F[s_A(x^{F,K}, Y)]\right| = 
% L\left| \sum_{i=1}^{N}\mathbb{E}_{F_i}\left[(Y_i-q_{\tilde{\tau},F_i})_+ - (Y_i-x_i^{F,K})_+\right]\right|
% \end{align}
where $\bm{q}_{\tilde{\tau},F}$ solves the linear program
$\sum_{x \in \undertilde{\bm{F}}^{-1}(\tilde{\tau})}x_i = K$.
